Formalizing Stack Safety as a Security Property

### 2023-01-24 submission5/intro-v2.tex

\rb{\textbf{The call stack and its security.} Maybe a few opening words to set
  the stage and bring two things to the front: (1) functions and subroutines as
  fundamental building blocks of any kind of structured programming; and (2) the
  stack as the natural abstraction to manage these in a computer. And also maybe
  a little bit more about the width, the gravity and the \emph{foundational
  nature} of attacks on the stack, which the sentence below touches on
  summarily.}
%
The call stack is a perennial target for low-level attacks, leading to
dire consequences from leakage and corruption of private stack data
to control-flow hijacking.
%
\rb{We can mention things like e.g. the most recent CWE Top 25 Most Dangerous
    Software Weaknesses from MITRE, cf.
    \url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}, where
    e.g. \#1 (Out-of-bounds Write), \#5 (Out-of-bounds Read) and \#7 (Use After
    Free), that is a good number of the most important and prevalent
    vulnerability types are closely related to the stack. These will show up in
    other popular rankings as well.}
%
To foil such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
  \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
\rb{Did we want to add even more techniques? Is this selection broad enough?
  Anyway reference that we'll be looking at these more closely later.}

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more formal characterization to the reader's intuition.
%
\rb{Begins to anticipate subsequent discussion. This is even something of an
  understatement: some papers (esp. newer ones?), make concrete claims, though
  they remain informal. Arguably, none have a truly formal characterization in
  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
  of the novelties of this work!}
%
Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
But this approach is not suitable for the {\em lazy} enforcement mechanism
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.
%
\rb{Need to discuss overlay semantics, but also note that it is very intensional
  and very specific, as the authors acknowledge. I'd introduce this a bit more
  gradually. Also, if phrased this way, it sounds like its main drawback is a
  lack of generality, to which we should contrast a more general approach, not
  just one that is specialized to a specific policy, i.e. lazy micro-policies
  (I'd try to present these more gradually as well).}

@@

We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak uses the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.
%
\rb{Anything to add from personal communications? More details later? Shows
  these things are very hard to get right -- like the latest work on Cerise,
  also motivated by this very work.}

@@

@@

@@

>>>>

\rb{\textbf{The call stack and its security.} Maybe a few opening words to set
  the stage and bring two things to the front: (1) functions and subroutines as
  fundamental building blocks of any kind of structured programming; and (2) the
  stack as the natural abstraction to manage these in a computer. And also maybe
  a little bit more about the width, the gravity and the \emph{foundational
  nature} of attacks on the stack, which the sentence below touches on
  summarily.
\rb{(+SNA) Connection to functions (and adversarial code?). Attacks on the stack
  are also attacks on this fundamental abstraction (functions, subroutines,
  procedures, methods, however you call them)}
}
%
The call stack is a perennial target for low-level attacks, leading to
dire consequences from leakage and corruption of private stack data
to control-flow hijacking.
%
\rb{We can mention things like e.g. the most recent CWE Top 25 Most Dangerous
    Software Weaknesses from MITRE, cf.
    \url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}, where
    e.g. \#1 (Out-of-bounds Write), \#5 (Out-of-bounds Read) and \#7 (Use After
    Free), that is a good number of the most important and prevalent
    vulnerability types are closely related to the stack. These will show up in
    other popular rankings as well.
\rb{(+SNA) List major CWEs? No matter the class of weakness considered, many
  involve a \emph{flow} going through the stack (cf. Eternal war on memory).}
}
%
To foil such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges+21},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18}.
  \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
\rb{Did we want to add even more techniques? Is this selection broad enough?
  Anyway reference that we'll be looking at these more closely later.}

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more formal characterization to the reader's intuition.
%
\rb{Begins to anticipate subsequent discussion. This is even something of an
  understatement: some papers (esp. newer ones?), make concrete claims, though
  they remain informal. Arguably, none have a truly formal characterization in
  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
  of the novelties of this work!
\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
  than previous work, without showing a fully general framework (which would be
  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
  start because of their flexibility, elegance, enabling a wide of interesting
  experiments, etc.}
}
%
Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
But this approach is not suitable for the {\em lazy} enforcement mechanism
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.
%
\rb{Need to discuss overlay semantics, but also note that it is very intensional
  and very specific, as the authors acknowledge. I'd introduce this a bit more
  gradually. Also, if phrased this way, it sounds like its main drawback is a
  lack of generality, to which we should contrast a more general approach, not
  just one that is specialized to a specific policy, i.e. lazy micro-policies
  (I'd try to present these more gradually as well).}

@@

We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak uses the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.
%
\rb{Anything to add from personal communications? More details later? Shows
  these things are very hard to get right -- like the latest work on Cerise,
  also motivated by this very work.
\rb{(+SNA) At the very least, these models are subtle and getting them right,
  and describing them unambiguously, is hard without a framework like this.}
}

@@

  \rb{(+SNA) They are kind of lazy, aren't they?}

@@

\rb{(+SNA) Discuss the degree in which the various papers satisfy the
  properties\ldots and the degrees to which they \emph{aim to}. E.g. originally
  they don't talk about confidentiality. Do they have callee integrity? Maybe
  they aren't thinking about it? (This kind of discussion is a good motivation
  for the formal properties. Also related to the hierarchy. And: how to convey
  that the latest paper draws inspiration directly from this work?}

@@

\rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
\rb{(+SNA) A general multi-level extension of the framework would need to model
  multiple source languages. Among other things, the stack is the meeting place
  of all sorts of different language abstractions! One of the advantages of the
  single-level model is that it lets us be agnostic about the language. (On the
  other hand, it is likely that a single source and target mediated by the
  compiler suffice to model many existing techniques, coming e.g. from systems
  programming and C-like languages.}
